---
title: "Homework 01: Data visualization"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Load packages and data

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

```{r read-in-data}
set.seed(1)
diamonds_subset <- diamonds %>%
  filter(carat <= 2.5) %>%
  slice_sample(n = 1000)
```

# Exercise 1

There are 1000 rows and 10 columns in this dataset. Note you can also just
examine the documentation, so no code is necessary.

```{r ex-1}
dim(diamonds_subset)
```

# Exercise 2

Clarity indicates how clear the diamond is, ranking from I1 (worst) to IF 
(best). Color indicates the diamond's color, ranking from D (best) to J (worst).
We should keep the meaning of these variables in mind as we answer these 
questions.

# Exercise 3

```{r ex-3}
ggplot(data = diamonds_subset, 
       aes(x = carat, y = price)) + 
  geom_point() +
  labs(title = "Price vs carat for 1,000 diamonds",
       y = "Price (USD)", x = "Carat")
```

Based on the scatterplot, it appears that larger diamonds are associated with
higher prices, and as the size of the diamond increases there is more
variability in price.

# Exercise 4

```{r ex-4}
ggplot(data = diamonds_subset, 
       aes(x = carat, y = price, color = color)) + 
  geom_point() +
  labs(title = "Price vs carat for 1,000 diamonds",
       subtitle = "colored by diamond color",
       y = "Price (USD)", x = "Carat")
```

For constant diamond weight, better colors are associated with higher prices.

# Exercise 5

```{r ex-5}
ggplot(data = diamonds_subset, 
       aes(x = carat, y = price, color = color)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "Price vs carat for 1,000 diamonds",
       y = "Price (USD)", x = "Carat")
```

# Exercise 6

```{r ex-6}
ggplot(data = diamonds_subset, 
       aes(x = carat, y = price)) + 
  geom_point() + 
  facet_wrap( ~ clarity) +
  labs(title = "Price vs carat for 1,000 diamonds",
       subtitle = "faceted by clarity",
       x = "Carat",
       y = "Price (USD)")
```

# Exercise 7

```{r ex-7}
ggplot(data = diamonds_subset, 
       aes(x = color)) + 
  geom_bar() + 
    labs(title = "Bar plots of diamond color",
       x = "Color",
       y = "Count")
```

# Exercise 8

```{r ex-8}
ggplot(data = diamonds_subset, 
       aes(x = color, fill = cut)) + 
  geom_bar(position = "fill") +  
  labs(title = "Bar plots of diamond color",
       subtitle = "filled by cut",
       x = "Color",
       y = "Proportion")
```

# Exercise 9

This does not (and should not) work as price is a numeric variable.

# Exercise 10

```{r ex-10}
ggplot(data = diamonds_subset, 
       aes(x = color, y = price)) + 
  geom_boxplot() +
  labs(title = "Side-by-side boxplots of price by color",
       x = "Color", 
       y = "Price (USD)")

ggplot(data = diamonds_subset, 
       aes(x = color, y = price)) + 
  geom_violin() + 
  labs(title = "Side-by-side violinplots of price by color",
       x = "Color", 
       y = "Price (USD)")
```

The boxplots show the first and third quartile, median, and any outliers, while
the violin plot shows a rotated density function for each price. Violin plots
give a better sense of the distribution of price within each diamond color.

# Exercise 11

Answers will vary.