---
title: "Homework 04: Statistical inference"
author: ""
date: ""
output:
  pdf_document: default
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, 
                      warning = FALSE, comment = "#>", highlight = TRUE)
set.seed(37073)
```

# Packages

```{r packages}
library(tidyverse)
library(infer)
```

# Data

```{r load_data}
gifted <- read_csv("https://www.openintro.org/data/csv/gifted.csv")
```

# Exercises

## Exercise 1

337,690 interviews conducted Jan. 2, 2016, through Dec. 30, 2017

## Exericse 2

These are sample proportions from the samples taken in the respective years.
They are being reported as point estimates for the true unknown population
parameters.

## Exericse 3

Confidence interval: (0.113, 0.117)

We are 95% confident that the true proportion of adults nationwide with 
diabetes is captured by the interval (0.113, 0.117).

## Exericse 4

Confidence interval: (0.049, 0.119)

We are 95% confident that the true proportion of adults in Alaska with 
diabetes is captured by the interval (0.049, 0.119).

## Exericse 5

$$H_0: p = 0.90$$
$$H_A: p < 0.90$$

Let $\alpha = 0.05$.

```{r ex_5}
claims <- tibble(settled_status = rep(c("on time", "not on time"), 
                                      times = c(55, 20)))

claims %>% 
  prop_test(response = settled_status, p = 0.90, alternative = "less", z = TRUE,
            success = "on time")
```

Since the p-value is less than $\alpha$, we reject the null hypothesis. Hence,
we reject the company's claim that 90% of its claims are settled within
30 days.

## Exercise 6

```{r ex_6}
claims %>% 
  specify(response = settled_status, success = "on time") %>% 
  hypothesise(null = "point", p = 0.90) %>% 
  generate(reps = 10000) %>% 
  calculate(stat = "prop") %>% 
  get_p_value(obs_stat = 55 / 75, direction = "less")
```

Both results are in agreement. Our conclusion does not change as the p-value
is less than $\alpha$.

## Exercise 7

```{r ex_7}
gifted %>% 
  specify(response = count) %>% 
  generate(reps = 10000) %>% 
  calculate(stat = "mean") %>% 
  get_ci(level = 0.98)
```

We are 98% confident that the mean number of months until gifted children from 
the specified city can count to 10 is captured by the interval (29.1, 32.3).

## Exercise 8 

```{r ex_8}
gifted %>% 
  pivot_longer(cols      = edutv:cartoons, 
               names_to  = "tv_type", 
               values_to = "tv_hours") %>%
  specify(tv_hours ~ tv_type) %>% 
  generate(reps = 10000) %>%
  calculate(stat = "diff in means", order = c("cartoons", "edutv")) %>% 
  get_ci(level = 0.99)
```

Since our confidence interval does not cover 0, we can reject any claim that
these children watch, on average, the same amount of cartoons and educational
tv at the $\alpha = 0.01$ significance level. 

## Exercise 9

Ranking from largest to smallest:

1. C
2. B
3. A

B's interval is wider than A's because they use the same data, but B uses a 
larger confidence level - `r qnorm(.975)` > `r qnorm(0.95)`. C's interval
has to be wider than B's because C used a smaller sample size and had
$\hat{p} = 0.50$. A smaller sample size increases the margin of error, all
else held constant. Also, for a fixed sample size, when $\hat{p} = 0.5$ the
standard error is maximized. Therefore, regardless of what value $\hat{p}$
B had, C will have a wider interval.

```{r}
ggplot() +
  xlim(0, 1) +
  stat_function(fun = ~ sqrt(.x * (1 - .x) / 100), color = "blue") +
  labs(y = "Standard error", x = "Sample proportion",
       caption = "n is fixed to be 100") +
  theme_minimal()
```

## Exercise 10 

$$H_0: \mu = 2.00$$
$$H_A: \mu > 2.00$$

A Type I error occurs when the data leads us to reject $H_0$, but $H_0$ is
actually true. This type of error would result in the company modifying its
production pipeline and possibly produce catheters that are not 2mm in diameter.

A Type II occurs when the data leads us to not reject $H_0$, but $H_0$ is
actually false. This type of error would result in the company continuing
to produce catheters as they currently do. It could result in catheters having
incorrect diameters, which would cause complications for patients.

# References

Gallup, I. (2018). Diabetes Rates Rise in 18 States in Past Decade. 
Gallup.com. Retrieved from 
https://news.gallup.com/poll/243911/diabetes-rates-rise-states-past-decade.aspx
