---
title: "Homework #02: Data wangling"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Load Packages and data

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

```{r read-in-data, message = FALSE, warning = FALSE}
people <- read_csv("data/people.csv")
pitching <- read_csv("data/pitching.csv")
salaries <- read_csv("data/salaries.csv")
teams <- read_csv("data/teams.csv")
```

# Exercise 1

```{r ex-1}
team_stats <- teams %>% 
  filter(year_id %in% 2011:2016) %>%
  mutate(win_pct = wins / games ) %>%
  select(year_id, team_id, win_pct)
```

# Exercise 2

```{r ex-2}
team_spending <- pitching %>%
  inner_join(salaries, by = c("player_id", "year_id", "team_id")) %>%
  group_by(year_id, team_id) %>%
  summarize(pitching_salaries = sum(salary, na.rm = TRUE)) %>%
  ungroup()

team_spending
```

# Exercise 3

```{r ex-3}
team_stats %>%
  left_join(team_spending, by = c("year_id", "team_id")) %>%
  ggplot(aes(x = pitching_salaries / 1e6, y = win_pct)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + 
  labs(title = "Win percentage versus pitching salaries",
       x = "Pitching Spending (millions)", y = "Win Percentage") + 
  facet_wrap(~ year_id) + 
  theme_bw()
```

# Exercise 4

```{r ex-4}
top_so_by_year <- pitching %>% 
  filter(year_id >= 1901) %>%
  inner_join(people, by = "player_id") %>%
  group_by(year_id, league_id) %>%
  filter(strike_outs == max(strike_outs)) %>%
  select(year_id, league_id, strike_outs, first_name, last_name)

glimpse(top_so_by_year)
```

# Exercise 5

```{r ex-5}
player_era <- pitching %>%
  group_by(player_id, year_id) %>%
  summarize(total_earned_runs = sum(earned_runs, na.rm = TRUE),
            total_strike_outs = sum(strike_outs, na.rm = TRUE),
            total_innings_pitched = sum(innings_pitched, na.rm = TRUE),
            era = 9 * (total_earned_runs / total_innings_pitched)) %>%
  select(player_id, year_id, total_strike_outs, era) %>%
  ungroup()
```

# Exercise 6

```{r ex-6}
player_era_age <- player_era %>%
  left_join(people, by = "player_id") %>%
  mutate(age = year_id - birth_year)

pitching_stats_by_age <- player_era_age %>%
  group_by(age) %>%
  summarize(median_era = median(era, na.rm = TRUE),
            median_so = median(total_strike_outs, na.rm = TRUE),
            count = n())
```

# Exercise 7

```{r ex-7}
ggplot(data = pitching_stats_by_age, 
       mapping = aes(x = age, y = median_so, size = count)) + 
  geom_point(alpha = 0.50) +
  labs(title = "Strike out performance as pitchers age",
       x = "Player age", y = "Median strikeouts", size = "Number of players") +
  theme_bw()
```

Higher player ages are associated with a higher median number of strikeouts for
players between the ages of 16 and 30. Between 30 and 35 or so, the median 
number of strikeouts is roughly stable. For pitchers older than 35, higher 
ages are associated with a lower median number of strikeouts, and then the 
median becomes highly variable, with older pitchers having both an 
exceptionally high and an exceptionally low number of strikeouts. This 
variability is a consequence of the small number of older pitchers.

Interestingly, there are four pitchers who are only 16 years old. Two of them,
Joe Nuxhall and Carl Scheib, pitched in the 20th century.

# Exercise 8

```{r ex-8}
top_players <- c("koufasa01", "kershcl01", "gibsobo01", "clemero02",
                 "maddugr01", "youngcy01", "johnswa01", "johnsra05",
                 "mathech01")

player_era_age %>%
  filter(player_id %in% top_players) %>%
  mutate(full_name = paste0(last_name, ", ", first_name)) %>%
  ggplot(aes(x = age, y = era)) + 
  geom_point() + 
  geom_line() + 
  labs(title = "How do pitchers age?",
       subtitle = "trends in ERA for careers of top pitchers",
       x = "Age", y = "Earned Run Average") +
  facet_wrap( ~ full_name) +
  theme_bw()
```

These top nine pitchers had careers ranging from eleven years (Clayton
Kershaw and Sandy Koufax) to 23 years (Roger Clemens). All (unsurprisingly) have
relatively low earned run averages. For reference, in modern baseball, an 
earned run average between 2 and 3 is excellent, and an earned run average below
2 is exceptional. The all-time leader is Ed Walsh, with a career earned run 
average of 1.82.

There does not appear to be a common trend. Koufax started with a fairly high
earned run average and improved over the course of his career, and Walter 
Johnson started with a low earned run average which got worse over the course of 
his career. Roger Clemens is highly variable. Christy Mathewson and Cy Young
have consistently low earned run averages. Of note is Bob Gibson's 1968 
season, (arguably) the best performance by a pitcher in modern baseball with
an earned run average of 1.12 in about 300 innings. 
